---
title: "case2"
author: "Richard Gan"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: true
    highlight: tango
    toc: true
    toc_float: true
    theme: journal
    toc_depth: 5
    number_sections: true
  word_document:
    toc: true
    toc_depth: 5
  pdf_document:
    toc: true
    toc_depth: '5'
    bibliography: case1_12287423.bib
    csl: "apa-7th-edition.csl"
---

```{r setup, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
require(tufte)
require(knitr)
require(tinytex)
```

# Introduction
Sector: Transportation Services
Companies: MacroAsia Corporation (PSE:MAC) and International Container Terminal Services, Inc. (PSE:ICT)
Interpretation: Interpretation is commented on last line(s) of each code chunk 
Note: For MacroAsia (MAC) price was used to model. This is to provide further analysis since Log Returns model generates ARIMA(0,0,0). Stationary data may have caused it.

# MAC Background 
MacroAsia Corporation (MAC), through its subsidiaries and associates, is engaged in aviation-support businesses at the Ninoy Aquino International Airport (NAIA), Manila Domestic Airport, Mactan-Cebu International Airport, Kalibo International Airport, Davao International Airport, and the General Aviation Areas. The group provides aircraft maintenance, repairs and overhaul services, in-flight catering services, airport ground handling services, charter flight services, and operates a special economic zone at the NAIA. 

# ICT Background 
International Container Terminal Services, Inc. (ICT) was incorporated on December 24, 1987 to operate, manage and develop the Manila International Container Terminal (MICT), which handles international container cargo at the Port of Manila. ICT provide ancillary services such as storage, container stripping and stuffing, inspection, weighing and services for refrigerated containers or reefers, as well as roll-on/roll-off and anchorage services to non-containerized cargoes or general cargoes on a limited basis.

```{r Packages}
pacman::p_load(tseries,quantmod,forecast,timeSeries,FinTS,rugarch,dplyr,PerformanceAnalytics,readxl,xts,tsm,remotes,fpp3, ggplot2, urca, plotly, ggfortify)
remotes::install_github("KevinKotze/tsm")
```
# PART 1: Data Preparation

```{r MacroAsia Data}
# Load MacroAsia 
mdata <- read_excel('MAC.xlsx', sheet = 'Sheet 1')

# XTS Conversion
mdata <- data.frame(date = as.Date(mdata$date), value = mdata$close)
mdata <- xts(mdata$value, order.by = mdata$date)
mdata <- na.omit(mdata)
plot(mdata, main = 'MacroAsia Price')
```

```{r ICT Data}
idata <- read_excel('ICT.xlsx', sheet = 'Sheet 1')

# XTS Conversion
idata <- data.frame(date = as.Date(idata$date), value = idata$close)
idata <- xts(idata$value, order.by = idata$date)
plot(idata, main = 'ICT Price')
```

```{r MAC log returns, volatility clustering}
mdata_ret <- na.omit(diff(log(mdata)))
plot(mdata_ret,main = 'MacroAsia Log returns')
# volatility clustering around 2017 and 2020 hence ARIMA not sufficient model
```


```{r ICT log returns, volatility clustering}
idata_ret <- na.omit(diff(log(idata)))
plot(idata_ret, main = 'ICT Log returns')
# volatility clustering around 2008 and 2020 
```

# PART 2: ARIMA Model Identification
```{r MAC stationarity test}
adf.test(mdata_ret) # H0: TS is not stationary
# Hypothesis testing = Fail to Reject Null hypothesis | Data is not stationary

kpss.test(mdata_ret, null = "Level") # H0: TS is level stationary
# Hypothesis testing = Reject Null hypothesis | Data is stationary
```


```{r ICT stationary test}
adf.test(idata_ret) # H0: TS is not stationary
# Fail to REJECT NUll Hypothesis | Data is stationary

kpss.test(idata_ret, null = "Level") # TS is level stationary
# Reject null hypothesis as data is not level stationary
```

```{r MAC ACF and PCF}
par(mfrow = c(1,2))
Acf(mdata_ret, main = "MAC ACF of Differenced Data")
Pacf(mdata_ret, main = "MAC PACF of Differenced Data")
# There is significant spikes at different lags exceeding the bands and slight decay
# Significant spikes indicates it is not white noise. Meaning it will become ARIMA(0,0,0) and we cannot use ARIMA  to model
```
```{r ICT ACF and PCF}
par(mfrow = c(1,2))
Acf(idata_ret, main = 'ICT ACF of Differenced Data')
Pacf(idata_ret, main = 'ICT PACF of Differenced Data')
# Significant spikes to the bands are observed around 0 to 5 Lag
# Significant spikes indicates it is not white noise. Meaning it will become ARIMA(0,0,0) and we cannot use ARIMA  to model
```
# PART 3: ARIMA Model
```{r MAC ARIMA Model}
mac_arima_model<- auto.arima(mdata, trace = T)
summary(mac_arima_model)
# Price is used for ARIMA model for this case, since Log Returns gets ARIMA (0,0,0) model
# ARIMA Model(2,1,2) = AR(2), Differencing = 1, MA(2)
mac_arima_model
```

```{r ICT ARIMA Model}
ict_arima_model<- auto.arima(idata_ret, trace = T)
summary(ict_arima_model)
# ARIMA(2,0,1) with non-zero mean  = AR(2), Differencing = 0, MA(1)
ict_arima_model
```

# PART 4: Residual Analysis
```{r MAC Residuals}
par(mfrow = c(1,2))
Acf(mac_arima_model$residuals, main = "MAC ACF of ARIMA Residuals")
Pacf(mac_arima_model$residuals, main = "MAC PACF of ARIMA Residuals")
# Significant spikes indicates it is not white noise. Meaning it will become ARIMA(0,0,0) and we cannot use ARIMA  to model
```
```{r ICT Residuals}
par(mfrow = c(1,2))
Acf(ict_arima_model$residuals, main = "ICT ACF of ARIMA Residuals")
Pacf(ict_arima_model$residuals, main = "ICT PACF of ARIMA Residuals")
# Significant spikes indicates it is not white noise. Meaning it will become ARIMA(0,0,0) and we cannot use ARIMA  to model
```

```{r MAC Box and coef test}
# MAC Box Test
Box.test(mac_arima_model$residuals, type = "Ljung-Box") # H0 : No autocorrelation in the residuals
# Reject Null Hypothesis : There is no autocorrelation in the residuals

# MAC Coeftest
lmtest::coeftest(mac_arima_model)
# AR2 and MA2 are significant in our ARIMA(2,1,2) Model
```
```{r ICT Box and coef test}
# ICT Box Test
Box.test(ict_arima_model$residuals, type = "Ljung-Box") # H0 : No autocorrelation in the residuals
# Reject Null Hypothesis : There is no autocorrelation in the residuals 

# ICT Coeftest
lmtest::coeftest(ict_arima_model)
# AR2 and MA2 are significant in our ARIMA(2,0,1) Model
```

```{r MAC check significant}
mac_arima_model2 <- Arima(mdata, order = c(
  1,0,1 #2,1,2
  )) # AR(1) and MA(1) components are significant
summary(mac_arima_model2)
lmtest::coeftest(mac_arima_model2)
```
```{r ICT check significant}
ict_arima_model2 <- Arima(idata, order = c(1,0,1)) # AR(1) and MA(1) components are significant
summary(ict_arima_model2)
lmtest::coeftest(ict_arima_model2)
```
```{r MAC check residuals}
forecast::checkresiduals(mac_arima_model)
```

```{r ICT check residuals}
forecast::checkresiduals(ict_arima_model)
```
```{r MAC AIC and BIC} 
# Lower AIC = Better
AIC(mac_arima_model) # Better Model per AIC
AIC(mac_arima_model2)
BIC(mac_arima_model)
BIC(mac_arima_model2)
```
```{r ICT AIC and BIC}
# Lower AIC = Better
AIC(ict_arima_model) # Better Model per AIC
AIC(ict_arima_model2)
BIC(ict_arima_model)
BIC(ict_arima_model2)
```

```{r MAC ArchTest}
mac_arima_model_test <- FinTS::ArchTest(residuals(mac_arima_model), lags = 12) # H0: No ARCH effects
mac_arima_model_test 
# Fail to Reject Null Hypothesis | There is Arch effects hence MAC ARIMA Model is insufficient
# At Lag 36 we Reject Null
```

```{r ICT ArchTest}
ict_arima_model_test <- FinTS::ArchTest(residuals(ict_arima_model), lags = 12) # H0: No ARCH effects
ict_arima_model_test
# Reject Null Hypothesis | There is Arch effects hence ICT ARIMA Model is insufficient
# All Lags have ARCH effects
```

```{r Final ARIMA Model}
#mac_arima_model # ARIMA(2,1,2)
ict_arima_model # ARIMA(2,0,1)
# The ARIMA Models are not sufficient as they have ARCH effects however the model is significant
```

======================================================

# PART 5 : GARCH Model

```{r ICT sGARCH nd}
ict_garch_spec_nd <- ugarchspec(
  mean.model = list(
     armaOrder = c(2,1)                     # 2 AR terms, 1 MA term | Adjusted for ARMA(2,0,1) 
    ,include.mean = TRUE                    # explicitly include mean (mu)
    ,fixed.pars = list(ma2 = 0, ma3 = 0))   # Fix ma2 and ma3 to 0
  ,variance.model = list(
     model = "sGARCH"                       # GARCH model for volatility
    ,garchOrder = c(1, 1))                  # GARCH(1,1) model for conditional variance
  ,distribution.model = "norm"              # sGARCH Normal Distribution
)
ict_garch_model_nd <- ugarchfit(spec = ict_garch_spec_nd
                             ,data = ict_arima_model$residuals
                             )
ict_garch_model_nd
show(ict_garch_spec_nd)
coef(ict_garch_model_nd)

# Ljung-Box Test indicates model is not fit and there is autocorrelation 
# Squared residuals shows no autocorrelation
# There is no ARCH effects on residuals
# Joint statistic infers that our parameters are jointly unstable and individually the results are mixed and sensitive to changes
# Pearson Goodness-of-Fit infers model is not fit
# Sign bias shows the model is not able to capture positive and negative shocks
# Negative shocks vs positive shocks | Model appears to have a slight bias towards positive 
```

```{r MAC sGARCH nd}
mac_garch_spec_nd <- ugarchspec(
  mean.model = list(
     armaOrder = c(2,2)                     # 2 AR terms, 2 MA term | Adjusted for ARMA(2,1,2) 
    ,include.mean = TRUE                    # explicitly include mean (mu)
    ,fixed.pars = list(ma2 = 0, ma3 = 0))   # Fix ma2 and ma3 to 0
  ,variance.model = list(
     model = "sGARCH"                       # GARCH model for volatility
    ,garchOrder = c(1, 1))                  # GARCH(1,1) model for conditional variance
  ,distribution.model = "norm"              # sGARCH Normal Distribution
)
mac_garch_model_nd <- ugarchfit(spec = mac_garch_spec_nd
                             ,data = mac_arima_model$residuals
                             )
mac_garch_model_nd
show(mac_garch_spec_nd)
coef(mac_garch_model_nd)

# Ljung-Box Test indicates model is  fit and there is no autocorrelation 
# Squared residuals shows no autocorrelation
# There is no ARCH effects on residuals
# Joint statistic infers that our parameters are jointly stable and individually the results are mostly stable
# Pearson Goodness-of-Fit infers model is not fit
# There is no sign bias 
```

```{r sGARCH sstd}
ict_garch_spec_std <- ugarchspec(
  mean.model = list(
     armaOrder = c(2,1)                     # 2 AR terms, 1 MA term | Adjusted for ARMA(2,0,1) 
    ,include.mean = TRUE                    # explicitly include mean (mu)
    ,fixed.pars = list(ma2 = 0, ma3 = 0))   # Fix ma2 and ma3 to 0
  ,variance.model = list(
     model = "sGARCH"                       # GARCH model for volatility
    ,garchOrder = c(1, 1))                  # GARCH(1,1) model for conditional variance
  ,distribution.model = "std"              # sGARCH Student's T Distribution
)
ict_garch_model_std <- ugarchfit(spec = ict_garch_spec_std
                             ,data = ict_arima_model$residuals
                             )
ict_garch_model_std
show(ict_garch_spec_std)
coef(ict_garch_model_std)

# Higher Log likelihood = Better
# Lower AIC = Better
# Both models (norm vs std) show preferable results
```

```{r GJR-GARCH}
ict_gjrgarch_spec_std <- ugarchspec(
  mean.model = list(
     armaOrder = c(2,1)                     # 2 AR terms, 1 MA term | Adjusted for ARMA(2,0,1) 
    ,include.mean = TRUE                    
    ,fixed.pars = list(ma2 = 0, ma3 = 0))   
  ,variance.model = list(
     model = "gjrGARCH"                     # GJR GARCH model 
    ,garchOrder = c(1, 1))                
  ,distribution.model = "std"             
)
ict_gjrgarch_model_std <- ugarchfit(spec = ict_gjrgarch_spec_std
                             ,data = ict_arima_model$residuals
)
ict_gjrgarch_model_std
show(ict_gjrgarch_spec_std)
coef(ict_gjrgarch_model_std)
# Higher Log likelihood and AIC shows little change from sGARCH model
```
```{r eGARCH}
# Exponential GARCH
ict_egarch_spec <- ugarchspec(variance.model = list(model = "eGARCH", garchOrder = c(1, 1)),
                          mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
                          distribution.model = "std")
ict_egarch_model <- ugarchfit(spec = ict_egarch_spec, data = ict_arima_model$residuals)
ict_egarch_model
#show(ict_egarch_model)
# No material difference from previous sGARCH models
```

```{r TGARCH}
# Fit TGARCH model for comparison
ict_tgarch_spec <- ugarchspec(variance.model = list(model = "fGARCH", submodel = "TGARCH", garchOrder = c(1, 1)),
                          mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
                          distribution.model = "norm")
ict_tgarch_model <- ugarchfit(spec = ict_tgarch_spec, data = ict_arima_model$residuals)
ict_tgarch_model
# Log likelihood went down and AIC as well. Therefore, not much significant improvement from previous sGARCH models
```

# PART 6 : APARCH

```{r APARCH}
ict_aparch_spec <- ugarchspec(
  variance.model = list(model = "apARCH",  # APARCH model
                        garchOrder = c(1, 1)),  # GARCH(1,1)
  mean.model = list(armaOrder = c(2, 1), 
                    include.mean = TRUE), 
  fixed.pars = list(ma2 = 0, ma3 = 0),  # MA(0,4) with only ma1 and ma4
  distribution.model = "std"  # Standardized t-distribution
)

ict_aparch_model <- ugarchfit(spec = ict_aparch_spec, data = ict_arima_model$residuals)
ict_aparch_model
# There is no material difference in Log likelihood and AIC from the previous sGARCH,TGARCH,eGARCH Models
```

# Conclusion
MAC is not a viable ARIMA model as the data is stationary and there is no clear trend. Hence it returns ARIMA(0,0,0) however we have used price instead to get a viable analysis but this introduced a differencing factor ARIMA(2,1,2).

ICT was a viable and good model. We modeled ARIMA(2,0,1) and modeled volatility using GARCH. The results for the other GARCH was not material and sGARCH can be used to model the volatility sufficiently. 











